Incorporating Gyroscope Information in Gesture Recognition

### 2011-12-17 paper/paper.tex
table changes

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Features & Algorithm & Gestures Types & Accuracy \\ \hline
  Yes	&	Both  &	All & Cluster-matching with HMM (10) &	94.6\% \\ \hline
  Yes	&	Both	&	All & Original  &	89.6\% \\ \hline
  No	&	Both	&	All & Original  &	88.4\% \\ \hline
  Yes	&	Accel &	All & Cluster-matching with HMM (10) &	92.0\% \\ \hline
  Yes	&	Accel &	All &	Original & 84.2\% \\ \hline
  Yes	&	Gyro  &	All & Cluster-matching with HMM (10) & 92.3\% \\ \hline
  Yes	&	Gyro  &	All & Original &  89.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (10) &  96.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (20) &  97.9\% \\ \hline
  Yes	&	Both  &	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  No	&	Both	&	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  Yes	&	Both	& Squares &	Cluster-matching (10) & 	100\% \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching (10) & 	Chance \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching with HMM (10) &	72.0\% \\ \hline
  Yes	&	Both  & Circles and Reversed-circles  &	Cluster-matching with HMM (20) & 88.8\% \\
  \hline
\end{tabular}
\end{center}

@@

However, on the same data, our cluster-matching algorithm combined with
an HMM reaches 94.6\% accuracy.  We also observed similar improvements with
other examples, as seen in the data.

\subsection{HMM}
Although we found a few examples where the cluster-matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we found that adding the HMM actually decreased accuracy slightly,
from 96.7\% to 94.6\%, which may be due to random fluctuations.

However, when comparing the data for circles to our contrived data
set consisting of the first and second halves of each circle switched,
the using an HMM with alphabet size of 20 improved the results from chance
to 88.8\%.  Unfortunately we could not get such dramatic improvements with
any non contrived data.

@@

% For example, on the five original gestures, increasing the alphabet size
% from 10 to 40 improves the accuracy from 96.7\% to 98.4\%.  On the contrived
% circle data, the improvement was much larger, from 72.0\% to 88.8\%.

@@

\subsection{Normalization}
In general, our normalization makes only a small difference.  With the
original algorithm we saw an improvement of less than 1.5\% in accuracy,
and we saw no detectable difference using cluster-matching.

>>>>

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Features & Gestures Types & Algorithms & Accuracy \\ \hline
  Yes	&	Both  &	All & Cluster-Matching-HMM (10) &	94.6\% \\ \hline
  Yes	&	Both	&	All & KHMM  &	89.6\% \\ \hline
  No	&	Both	&	All & KHMM  &	88.4\% \\ \hline
  Yes	&	Accel &	All & Cluster-Matching-HMM (10) &	92.0\% \\ \hline
  Yes	&	Accel &	All &	KHMM & 84.2\% \\ \hline
  Yes	&	Gyro  &	All & Cluster-Matching-HMM (10) & 92.3\% \\ \hline
  Yes	&	Gyro  &	All & KHMM &  89.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-Matching (10) &  96.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-Matching (20) &  97.9\% \\ \hline
  Yes	&	Both  &	All &	Cluster-Matching (40) & 	98.4\% \\ \hline
  No	&	Both	&	All &	Cluster-Matching (40) & 	98.4\% \\ \hline
  Yes	&	Both	& Squares &	Cluster-Matching (10) & 	100\% \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-Matching (10) & 	Chance \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-Matching-HMM (10) &	72.0\% \\ \hline
  Yes	&	Both  & Circles and Reversed-circles  &	Cluster-Matching-HMM (20) & 88.8\% \\
  \hline
\end{tabular}
\end{center}

@@

However, on the same data, our Cluster-Matching algorithm combined with
an HMM reaches 94.6\% accuracy.  We also observed similar improvements with
other examples, as seen in the data.

\subsection{HMM}
Although we found a few examples where the Cluster-Matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we found that adding the HMM actually decreased accuracy slightly,
from 96.7\% to 94.6\%, which may be due to random fluctuations.

However, when comparing the data for circles to our contrived data
set consisting of the first and second halves of each circle switched,
then using an HMM with alphabet size of 20 improved the results from chance
to 88.8\%.  Unfortunately we could not get such dramatic improvements with
any non contrived data.

@@

% For example, on the five KHMM gestures, increasing the alphabet size
% from 10 to 40 improves the accuracy from 96.7\% to 98.4\%.  On the contrived
% circle data, the improvement was much larger, from 72.0\% to 88.8\%.

@@

\subsection{Normalization}
In general, our normalization makes only a small difference.  With the
KHMM algorithm we saw an improvement of less than 1.5\% in accuracy,
and we saw no detectable difference using Cluster-Matching.
